---
title: "EUSILC"
execute:
  warning: false
---

## Definition

EU-SILC is an EU statistics on income and living conditions. According to eurostat, EU-SILC is to gather up-to-date and comparable data on income, poverty, social exclusion, and living conditions, both in a cross-sectional and longitudinal manner.

## Data types

EU-SILC offers two kinds of data:

1.  Cross-sectional data, which captures information at a specific time or over a defined period, includes variables related to income, poverty, social exclusion, and various living conditions.
2.  Longitudinal data, which tracks changes at the individual level across time, observed regularly over four years.

Data on social exclusion and housing conditions are primarily gathered at the household level. Meanwhile, information on labor, education, and health is collected from individuals aged 16 and older. Detailed component-level income variables are also predominantly obtained from individual respondents.

## Data availablity

Data is available at eurostat database under the folder ["Living conditions and welfare"](https://ec.europa.eu/eurostat/web/main/data/database?node_code=livcon).

The folder contains four sub folders containing different datasets.

```         
1.  Consumption expenditure of private households (hbs)

    1.1 Mean consumption expenditure of private households (hbs_exp) 
    1.2 Structure of mean consumption expenditure (hbs_struc) 
    1.3 Household characteristics (hbs_carac)

2. Income and living conditions (ilc) 
    
    2.1 Persons at risk of poverty or social exclusion (EU 2030 target) (ilc_pe)

         2.1.1 Main indicator - Persons at risk of poverty or social exclusion (ilc_peps)
         2.1.2 Intersections between sub-populations of indicators included in the EU 2030 target on poverty and social exclusion (ilc_pees)
         
    2.2 Persons at risk of poverty or social exclusion and intersections between sub-populations (EU 2020 strategy) (ilc_p)
    
    2.3 Inequality (ilc_ie)
        2.3.1 Income inequality (ilc_iei)
        
    2.4 Income distribution and monetary poverty (ilc_ip)
        
        2.4.1 Monetary poverty (ilc_li)
        2.4.2 Monetary poverty for elderly people (ilc_pn)
        2.4.3 In-work poverty (ilc_iw)
        2.4.4 Distribution of income (ilc_di)
        
    2.5 Living conditions (ilc_lv)
    
        2.5.1 Private households (ilc_lvph) 
        2.5.2 Population structure (ilc_lvps) 
        2.5.3 Health and labour conditions (ilc_lvhl) 
        2.5.4 Housing conditions (ilc_lvho) 
        2.5.5 Childcare arrangements (ilc_ca) 
        2.5.6 Subjective poverty (ilc_sbjp) 
        
    2.6 Material deprivation (ilc_md)
    
        2.6.1 Material and social deprivation (ilc_mdsd)
        2.6.2 Material deprivation by dimension (ilc_mddd)
        2.6.3 Child specific material deprivation (ilc_chmd)
        2.6.4 Economic strain (ilc_mdes)
        2.6.5 Economic strain linked to dwelling (ilc_mded)
        2.6.6 Durables (ilc_mddu) 
        2.6.7 Housing deprivation (ilc_mdho) 
        2.6.8 Environment of the dwelling (ilc_mddw) 
        
    2.7 EU-SILC modules (ilc_ahm) 
    
        2.7.1 Living arrangements and conditions of children in separated and blended families (2021) (ilc_chsb) 
        2.7.2 Over-indebtedness, consumption and wealth as well as labour (2020) (ilc_odbt) 
        2.7.3 Intergenerational transmission of disadvantages (2019 and 2011) (ilc_igt) 
        2.7.4 Personal well-being indicators (2018 and 2013) (ilc_pwb)
        2.7.5 Health including children health (2017, 2021 and 2022) (ilc_hch)
        2.7.6 Access to services (2016) (ilc_ats)
        2.7.7 Social and cultural participation (2015) (ilc_scp)
        2.7.8 Material deprivation (2014) (ilc_mdm)
        2.7.9 Housing conditions (2012) (ilc_hcm)

3. Time use survey (tus)
4. Gender based violence against women (gbv)
```

To retrieve data from eurostat the following code can be used

```{r}
#| label = "load-pkg-data"

library(eurostat)
EU_SILC <- search_eurostat("EU-SILC survey")
knitr::kable(head(EU_SILC))

```


```{r}
# Calculate the amount of missing data in each column
library(dplyr)
missing_data <- sapply(EU_SILC, function(x) sum(is.na(x)))
missing_data
```


```{r}
#| results = "asis"
library(ggplot2)

# Function to create histograms for all numeric columns
plot_histograms <- function(EU_SILC) {
  numeric_columns <- sapply(EU_SILC, is.numeric)
  for (col_name in names(EU_SILC)[numeric_columns]) {
    p <- ggplot(EU_SILC, aes_string(x = col_name)) +
      geom_histogram(bins = 30, fill = "blue", color = "black") +
      theme_minimal() +
      labs(title = paste("Histogram of", col_name))
    print(p)
  }
}

# Call the function
plot_histograms(EU_SILC)
```

```{r}
#| results = "asis"
# Get the id from the table
id <- EU_SILC$code[1]

# Check the id
dat <- get_eurostat(id, time_format = "num", stringsAsFactors = TRUE)
knitr::kable(head(dat))
```



The [data availability table](https://ec.europa.eu/eurostat/documents/203647/771732/Datasets-availability-table.pdf) shows the latest and planned releases of EU-SILC scientific use files.

The [EU-SILC Guidelines ](https://ec.europa.eu/eurostat/documents/203647/16993001/Methodological+guidelines+2022+operation+v7.pdf/ec6bc779-6462-34aa-a8bd-256f8af34d31?t=1703153300474) can be refered for further information.

[List of EU-SILC microdata releases with digital object identifiers](https://ec.europa.eu/eurostat/documents/203647/203704/EU-SILC+microdata+with+DOIs.pdf/04b57b6d-803b-fa8d-bf6b-fe21fe6d1cb2?t=1703162264564)

## retroharmonize implementation

```{r}
#| results = "asis"
library(retroharmonize)
(attributes(dat))$names
```

## References

[Online Publications](https://ec.europa.eu/eurostat/statistics-explained/index.php?title=Online_publications)
